Supplements for the manuscript `Variation in gaze understanding across the life span: A process-level perspective'

### 2023-10-04 stats/gazemodel.Rmd
ready for first round of co-author feedback

# careful, this takes a while... 
model_kids <- readRDS("../data/gazemodel-kids.rds")
model_kids_dev_prior <- readRDS("../data/gazemodel-kids-dev-prior.rds")
```

@@

# CHILDREN

## Inference

```{r plot2B}
# (plot 2A is model logic, png)
# developmental trajectory model inference parameter
inf_map_devp <- model_kids_dev_prior %>%
  filter(scope == "global", 
         parameter == "inference", 
         type != "sigma") %>%
  group_by(type) %>%
  summarise(mode = estimate_mode(value)) %>%
  pivot_wider(names_from = "type", values_from = "mode") %>%
  expand_grid(age = seq(0, 3, 0.1)) %>%
  mutate(sd = exp(intercept + slope * age))

@@

dev_plot <- ggplot() +
  geom_point(data = id_devp, aes(x = age, y = mode), alpha = 0.75, color = "darkgrey") +
  geom_line(data = inf_samp_devp, aes(x = age + 3, y = sd, group = paste(chain,iteration)), col = "blue", alpha = .01) +
  geom_line(data = inf_map_devp, aes(x = age + 3, y = sd), col = "black", linewidth = 1.25) +
  scale_x_continuous(limits = c(3, 6), breaks = seq(3, 6, 0.5)) + 
  scale_y_continuous(limits = c(0, 6), breaks = seq(0, 6, 0.25)) +
  coord_cartesian(ylim = c(0, 3)) +
  labs(x = "Age in years", y = expression(paste("Mode model parameter (", italic("\U03C3"[v]), ")")))

@@

```{r plot2C}
comparison_model_data_devp <- model_kids_dev_prior %>%
  filter(scope == "id", type == "sd") %>%
  group_by(id) %>%
  summarise(model_mode = estimate_mode(value)) %>%
  left_join(
    tango_kids %>% 
      group_by(subj_id, age, age_group) %>%
      summarise(data_mean = mean(click_dist_balloons)) %>%
      rename(id = subj_id)
  )

model_vs_data_plot <- ggplot(comparison_model_data_devp, aes(x = model_mode, y = data_mean)) +
  geom_point(aes(col = age)) +
  stat_cor(method = "pearson", label.x = 0.05, label.y = 4.9, aes(label = paste(..r.label..)), size = 4, r.accuracy = 0.01, cor.coef.name = "r") +
  geom_smooth(method = "lm", col = "darkgrey", alpha = 0.25) +
  scale_color_viridis_c(direction = -1, alpha = 0.9, limits = c(3, 6), breaks = seq(3, 6, 0.5)) +
  labs(x = expression(paste("Mode model parameter (", italic("\U03C3"[v]), ")")), y = "Mean data imprecision", color = "Age") + 
  scale_x_continuous(limits = c(0, 2.1), breaks = seq(0, 2, 0.5)) + 
  scale_y_continuous(limits = c(0, 5), breaks = seq(0, 5, 1))

model_vs_data_plot
saveRDS(model_vs_data_plot, file = "../saves/gazemodel_model_vs_data.rds")

@@

## Inference parameters individuals

@@

## Recover U pattern

@@

```{r plot2E}
# (plot 2D is model geometry logic, png)
# u pattern recovery
ind_sd <- model_kids_dev_prior %>%
  filter(scope == "id", type == "sd") %>%
  group_by(id) %>%
  summarise(sdTol = estimate_mode(value)) %>%
  left_join(tango_kids %>% rename(id = subj_id,
                          eyeCenterLY = eye_center_left_y,
                          eyeCenterLX = eye_center_left_x,
                          eyeCenterRY = eye_center_right_y,
                          eyeCenterRX = eye_center_right_x,
                          pupilLX = pupil_final_left_x,
                          pupilRX = pupil_final_right_x,
                          targetCenterX = target_center_x,
                          targetCenterY = target_center_y) %>%
              mutate(clickDist = abs(click_dist_from_target_center_x)) %>%
              select(id, age, age_group, trial_nr, eyeCenterLX, eyeCenterLY, eyeCenterRX, eyeCenterRY, pupilLX, pupilRX, targetCenterY, targetCenterX, clickDist))

@@

# saveRDS(ind_sd_long, file = "../saves/gazemodel_ind_sd_long.rds")

# priorMode <- model_kids_dev_prior %>%
#   filter(scope == "global",
#          parameter == "priorSd",
#          type != "sigma") %>%
#   group_by(type) %>%
#   summarise(mode = estimate_mode(value))
#

@@

u_plot <- modelpred_gaze %>%
  pivot_longer(cols = c(data_mean, model_mean), names_to = "source", values_to = "imprecision") %>%
  mutate(
    age_group = factor(age_group, levels = c(3, 4, 5), labels = c("3-year-olds", "4-year-olds", "5-year-olds")), 
  ) %>% 
  ggplot(aes(x = targetBin, y = imprecision, col = source)) +
  geom_point(alpha = 0.8) +
  facet_grid(~ age_group) +
  scale_y_continuous(limits = c(0, 5.5), breaks = seq(0, 5.5, 0.5)) +
  scale_color_manual(values = c("darkgrey", "darkblue"), labels = c("Data", "Model prediction")) +
  labs(x = "Target bin", y = "Imprecision", color = "Source") + 
  theme_few() +
  theme(
    legend.position = c(0.84 , 0.8), # c(0,0) bottom left, c(1,1) top-right
    strip.text = element_text(face = "bold", size = 12), 
    legend.title = element_text(size = 6),
    legend.text = element_text(size = 6), 
    legend.box.background = element_rect(), 
    legend.key.height = unit(0, "cm"),
    legend.key.width = unit(-1.5, "cm"),
    plot.margin = margin(0.2, 0, 0.1, 0.2, "cm") # t r b l
  )

@@

modelpred_gaze_plot
saveRDS(modelpred_gaze_plot, "../saves/gazemodel_modelpred_gaze_plot.rds")
# ggsave("../figures/gazemodel_modelpred_gaze_plot.png", width = 8, height = 5, scale = 1, bg = "white")
# ggsave("../figures/gazemodel_modelpred_gaze_plot.pdf", width = 8, height = 5, scale = 1, bg = "white")
```
# Alternative models click generation

@@

```{r plot2F}
modelpred_gaze_plot <- readRDS("../saves/gazemodel_modelpred_gaze_plot.rds")
modelpred_center_plot <- readRDS("../saves/gazemodel_modelpred_center_plot.rds")
modelpred_random_plot <- readRDS("../saves/gazemodel_modelpred_random_plot.rds")

@@

  modelpred_gaze_plot + 
    theme(legend.position = c(0.85, 0.2), 
          legend.title = element_text(size = 6),
          legend.text = element_text(size = 6), 
          legend.box.background = element_rect(), 
          legend.key.height = unit(0, "cm"),
          legend.key.width = unit(-1.5, "cm"),
          plot.margin = margin(0.2, 0, 0.1, 0.2, "cm") # t r b l
         ),

@@

  ncol = 3, 
  nrow = 2, 
  widths = c(11, 1, 12), 
  labels = c("A", "", "B", "C", "", "D")
)

gazemodel_plot
saveRDS(gazemodel_plot, "../saves/gazemodel_plot.rds")
ggsave("../figures/gazemodel_plot.png", width = 8, height = 5, scale = 1.3, bg = "white")
ggsave("../figures/gazemodel_plot.pdf", width = 8, height = 5, scale = 1.3, bg = "white", device = cairo_pdf) 
# add cairo_pdf to enable greek symbols in pdf saving
```

```{r results_table}
# create tibble for reading out stats results in rmd manuscript
corr_gazemodel_param <- comparison_model_data_devp %>% select(model_mode, data_mean) %>% corr.test(method = "pearson")

corr_gazemodel_pred <- modelpred_gaze %>% ungroup() %>% select(model_mean, data_mean) %>% corr.test(method = "pearson")
corr_centermodel_pred <- modelpred_center %>% ungroup() %>% select(model_mean, data_mean) %>% corr.test(method = "pearson")
corr_randommodel_pred <- modelpred_random %>% ungroup() %>% select(model_mean, data_mean) %>% corr.test(method = "pearson")

corrs <- rbind(corr_gazemodel_param$ci, corr_gazemodel_pred$ci, corr_centermodel_pred$ci, corr_randommodel_pred$ci) %>% 
  rownames_to_column(var = "corr") %>% 
  mutate(
    corr = recode(
            corr,
            `mdl_m-dt_mn` = "gazemodel_param", 
            `mdl_m-dt_mn1` = "gazemodel_pred", 
            `mdl_m-dt_mn2` = "centermodel_pred", 
            `mdl_m-dt_mn3` = "randommodel_pred", 
           )
  ) %>% 
  mutate(across(where(is.numeric), round, 2))

@@

```{r read_data_mcomparison}
modelcomparison_kids_bf <- readRDS("../data/gazemodel-kids-comparison-bf.rds")
modelcomparison_kids_samples <- readRDS("../data/gazemodel-kids-comparison-samples.rds")
```
```{r mcomparison}
# same as reading in modelcomparison_kids_bf
modelcomparison_kids_samples %>%
  group_by(model) %>%
  summarise(margllh = logSumExp(like)) %>%
  pivot_wider(names_from = model, values_from = margllh) %>%
  summarise(bf_inf_center = inf - center, 
            bf_inf_guess = inf - guess)
```

>>>>

# careful, this takes a while... 
model_kids <- readRDS("../data/gazemodel-kids.rds")
model_kids_dev_prior <- readRDS("../data/gazemodel-kids-dev-prior.rds")

model_adults_ind <- readRDS("../data/gazemodel-adults-ind.rds") %>% 
  rename("mode" = sdTol) %>% 
  mutate(
    age_group = 18, 
    age = 18,
    id = str_replace_all(id, "gafo_prolific_modelvalidation_vp", "gp1-")
  )

model_adults_bins <- readRDS("../data/gazemodel-adults-bins.rds") %>% 
  mutate(
    age_group = 18, 
    # transform into target width units
    data_mean = data_mean / 160, 
    model_mean = model_mean / 160,
  )
```

@@

# Inference

```{r plot3A}
# developmental trajectory model inference parameter
inf_map_devp <- model_kids_dev_prior %>%
  filter(scope == "global", 
         parameter == "inference", 
         type != "sigma") %>%
  group_by(type) %>%
  summarise(mode = estimate_mode(value)) %>%
  pivot_wider(names_from = "type", values_from = "mode") %>%
  expand_grid(age = seq(0, 3, 0.1)) %>%
  mutate(sd = exp(intercept + slope * age))

@@

model_adults_sd <- model_adults_ind %>% 
  summarize(
    mean_mode = mean(mode), 
    lci = hdi_lower(mode), 
    uci = hdi_upper(mode)
  )

dev_plot <- ggplot() +

  # children
  geom_point(data = id_devp, aes(x = age, y = mode), alpha = 0.75, color = "darkgrey") +
  geom_line(data = inf_samp_devp, aes(x = age + 3, y = sd, group = paste(chain,iteration)), col = "blue", alpha = .01) +
  geom_line(data = inf_map_devp, aes(x = age + 3, y = sd), col = "black", linewidth = 1.25) +

  # adults 
  geom_jitter(data = model_adults_ind, aes(x = 6.5, y = mode), alpha = 0.5, color = "darkgrey",  width = .1) + 
  geom_pointrange(data = model_adults_sd %>% head(1), aes(x = 6.5, y = mean_mode, ymin = lci, ymax = uci), size = 0.5) +

  scale_x_continuous(limits = c(3, 6.7), breaks = seq(3, 6.5, 0.5), labels = c("3.0", "3.5", "4.0", "4.5", "5.0", "5.5", "6.0", "Adults")) + 
  scale_y_continuous(limits = c(0, 6), breaks = seq(0, 6, 0.25)) +
  coord_cartesian(ylim = c(0, 3)) +
  labs(x = "Age in years", y = expression(paste("Mode model parameter (", italic("\U03C3"[v]), ")")))

@@

```{r plot3B}
comparison_model_data_devp <- model_kids_dev_prior %>%
  filter(scope == "id", type == "sd") %>%
  group_by(id) %>%
  summarise(model_mode = estimate_mode(value)) %>%
  left_join(
    tango_kids %>% 
      group_by(subj_id, age, age_group) %>%
      summarise(data_mean = mean(click_dist_balloons)) %>%
      rename(id = subj_id)
  )

comparison_model_data_adults <- tango_adults %>% 
  group_by(subj_id) %>% 
  summarize(
    data_mean = mean(click_dist_balloons)
  ) %>% 
  left_join(model_adults_ind %>% select(-lci, -uci) %>% rename("model_mode" = mode, "subj_id" = id))

comparison_model_data <- full_join(comparison_model_data_devp, comparison_model_data_adults) %>% 
  mutate(sample = ifelse(age < 18, "children", "adults"))

model_vs_data_plot <- ggplot(comparison_model_data, aes(x = model_mode, y = data_mean, group = sample)) +
  geom_point(aes(col = age)) +
  stat_cor(data = comparison_model_data %>% filter(sample == "children"), aes(label = paste(..r.label..)), method = "pearson", size = 4, r.accuracy = 0.01, cor.coef.name = "r", label.x = 0, label.y = 4.5) +

  stat_cor(data = comparison_model_data %>% filter(sample == "adults"), aes(label = paste(..r.label..)), method = "pearson", size = 4, r.accuracy = 0.01, cor.coef.name = "r", label.x = 0, label.y = 2.25) +

  geom_smooth(method = "lm", col = "darkgrey", alpha = 0.25) +

  labs(x = expression(paste("Mode model parameter (", italic("\U03C3"[v]), ")")), y = "Mean data imprecision", color = "Age") +

  scale_x_continuous(limits = c(0, NA), n.breaks = 7) +
  scale_y_continuous(limits = c(0, NA), n.breaks = 5) +
  scale_color_viridis_c(direction = -1, alpha = 0.9, limits = c(3, 6), breaks = seq(3, 6, 0.5)) +
  theme_few() +
  theme(
    legend.position = c(0.37, 0.3),
    strip.text = element_text(face = "bold", size = 10), 
    legend.title = element_text(size = 8),
    legend.text = element_text(size = 8), 
    legend.box.background = element_rect(), 
    legend.key.height = unit(0.5, "cm"),
  ) +
  facet_wrap(~ factor(sample, levels = c("children", "adults"), labels = c("Child sample", "Adult sample")), scales = "free")

model_vs_data_plot

saveRDS(model_vs_data_plot, file = "../saves/gazemodel_model_vs_data.rds")

@@

# Inference parameters individuals

@@

# Recover U pattern

@@

```{r plot3C}
# u pattern recovery
ind_sd <- model_kids_dev_prior %>%
  filter(scope == "id", type == "sd") %>%
  group_by(id) %>%
  summarise(sdTol = estimate_mode(value)) %>%
  left_join(tango_kids %>% rename(id = subj_id,
                          eyeCenterLY = eye_center_left_y,
                          eyeCenterLX = eye_center_left_x,
                          eyeCenterRY = eye_center_right_y,
                          eyeCenterRX = eye_center_right_x,
                          pupilLX = pupil_final_left_x,
                          pupilRX = pupil_final_right_x,
                          targetCenterX = target_center_x,
                          targetCenterY = target_center_y) %>%
              mutate(clickDist = abs(click_dist_from_target_center_x)) %>%
              select(id, age, age_group, trial_nr, eyeCenterLX, eyeCenterLY, eyeCenterRX, eyeCenterRY, pupilLX, pupilRX, targetCenterY, targetCenterX, clickDist))

@@

# saveRDS(ind_sd_long, file = "../saves/gazemodel_ind_sd_long.rds")
# ind_sd_long <- readRDS("../saves/gazemodel_ind_sd_long.rds")

priorMode <- model_kids_dev_prior %>%
  filter(scope == "global",
         parameter == "priorSd",
         type != "sigma") %>%
  group_by(type) %>%
  summarise(
    mode = estimate_mode(value), 
    lci = hdi_lower(value), 
    uci = hdi_upper(value))

saveRDS(priorMode, file = "../saves/gazemodel_prior.rds")

@@

u_data <- full_join(modelpred_gaze, model_adults_bins)
saveRDS(u_data, "../saves/gazemodel_u_data.rds")

u_plot <- u_data %>%
  pivot_longer(cols = c(data_mean, model_mean), names_to = "source", values_to = "imprecision") %>%
  mutate(
    age_group = factor(age_group, levels = c(3, 4, 5, 18), labels = c("3-year-olds", "4-year-olds", "5-year-olds", "Adults")), 
  ) %>% 
  ggplot(aes(x = targetBin, y = imprecision, col = source)) +
  geom_point(alpha = 0.8) +
  facet_grid(~ age_group) +
  scale_y_continuous(limits = c(0, 5.5), breaks = seq(0, 5.5, 0.5)) +
  scale_color_manual(values = c("darkgrey", "darkblue"), labels = c("Data", "Model prediction")) +
  labs(x = "Target bin", y = "Imprecision", color = "Source") + 
  theme_few() +
  theme(
    legend.position = c(0.881, 0.7), # c(0,0) bottom left, c(1,1) top-right
    strip.text = element_text(face = "bold", size = 10), 
    legend.title = element_text(size = 8),
    legend.text = element_text(size = 8), 
    legend.box.background = element_rect(), 
    legend.key.height = unit(0.4, "cm"),
    legend.key.width = unit(-1.2, "cm"),
    plot.margin = margin(0.2, 0, 0.1, 0.2, "cm") # t r b l
  )

@@

modelpred_gaze_plot
saveRDS(modelpred_gaze_plot, "../saves/gazemodel_modelpred_gaze_plot.rds")
# ggsave("../figures/gazemodel_modelpred_gaze_plot.png", width = 8, height = 5, scale = 1, bg = "white")
# ggsave("../figures/gazemodel_modelpred_gaze_plot.pdf", width = 8, height = 5, scale = 1, bg = "white")
```

# Alternative models click generation

@@

```{r plot3D}
modelpred_gaze_plot <- readRDS("../saves/gazemodel_modelpred_gaze_plot.rds")
modelpred_center_plot <- readRDS("../saves/gazemodel_modelpred_center_plot.rds")
modelpred_random_plot <- readRDS("../saves/gazemodel_modelpred_random_plot.rds")

@@

  modelpred_gaze_plot + 
    theme(legend.position = c(0.85, 0.3), 
          legend.title = element_text(size = 8),
          legend.text = element_text(size = 8), 
          legend.box.background = element_rect(), 
          legend.key.height = unit(0.3, "cm"),
          legend.key.width = unit(-1.5, "cm"),
          plot.margin = margin(0.2, -2, 0.1, 0.2, "cm") # t r b l
         ),

@@

  ncol = 3, 
  nrow = 2, 
  widths = c(13, 1, 12), 
  labels = c("A", "", "B", "C", "", "D")
)

gazemodel_plot
saveRDS(gazemodel_plot, "../saves/gazemodel_plot.rds")
ggsave("../figures/gazemodel_plot.png", width = 8, height = 5, scale = 1.5, bg = "white")
ggsave("../figures/gazemodel_plot.pdf", width = 8, height = 5, scale = 1.4, bg = "white", device = cairo_pdf)
# add cairo_pdf to enable greek symbols in pdf saving
```

```{r results_table}
# create tibble for reading out stats results in rmd manuscript
corr_gazemodel_param_children <- comparison_model_data_devp %>% select(model_mode, data_mean) %>% corr.test(method = "pearson")

corr_gazemodel_param_adults <- comparison_model_data_adults %>% select(model_mode, data_mean) %>% corr.test(method = "pearson")

# this is children only 
corr_gazemodel_pred <- modelpred_gaze %>% ungroup() %>% select(model_mean, data_mean) %>% corr.test(method = "pearson")
corr_centermodel_pred <- modelpred_center %>% ungroup() %>% select(model_mean, data_mean) %>% corr.test(method = "pearson")
corr_randommodel_pred <- modelpred_random %>% ungroup() %>% select(model_mean, data_mean) %>% corr.test(method = "pearson")

corrs <- rbind(corr_gazemodel_param_children$ci, corr_gazemodel_param_adults$ci, corr_gazemodel_pred$ci, corr_centermodel_pred$ci, corr_randommodel_pred$ci) %>% 
  rownames_to_column(var = "corr") %>% 
  mutate(
    corr = recode(
            corr,
            `mdl_m-dt_mn` = "gazemodel_param_kids", 
            `mdl_m-dt_mn1` = "gazemodel_param_adults", 
            `mdl_m-dt_mn2` = "gazemodel_pred_kids", 
            `mdl_m-dt_mn3` = "centermodel_pred_kids", 
            `mdl_m-dt_mn4` = "randommodel_pred_kids", 
           )
  ) %>% 
  mutate(across(where(is.numeric), round, 2))

@@

```{r read_data_mcomparison}
modelcomparison_kids_bf <- readRDS("../data/gazemodel-kids-comparison-bf.rds")
modelcomparison_kids_samples <- readRDS("../data/gazemodel-kids-comparison-samples.rds")
```

```{r mcomparison}
# same as reading in modelcomparison_kids_bf
modelcomparison_kids_samples %>%
  group_by(model) %>%
  summarise(margllh = logSumExp(like)) %>%
  pivot_wider(names_from = model, values_from = margllh) %>%
  summarise(bf_inf_center = inf - center, 
            bf_inf_guess = inf - guess)
```

### 2023-10-04 stats/supplements_gaussianprocess.Rmd
ready for first round of co-author feedback

write.csv(mgp_triallevel, "../saves/mgp_triallevel.csv", quote = F, row.names = F)
saveRDS(mgp_triallevel, "../saves/mgp_triallevel.rds")
```

@@

write.csv(mgp_aggregated, "../saves/mgp_aggregated.csv", quote = F, row.names = F)
saveRDS(mgp_aggregated, "../saves/mgp_aggregated.rds")
```

@@

ggplot() +

  geom_smooth(
    data = pred_combined, 
    aes(x = age, y = Estimate, ymin = Q2.5, ymax = Q97.5, group = model, col = model), 
    stat = "identity", size = 0.5, alpha = 0.15
  ) 
```

```{r aggregated_plot}
ggplot() +

  geom_jitter(
    data = mdata_aggregated, 
    aes(x = age, y = mean_imprecision), 
    col = "darkgrey", alpha = 0.75
  ) +

  geom_smooth(
    data = pred_aggregated, 
    aes(x = age, y = Estimate, ymin = Q2.5, ymax = Q97.5), 
    stat = "identity", size = 1, alpha = 0.15, fill = "#006C66", col = "#006C66",
  ) +

  scale_x_continuous(limits = c(0, 80), breaks = seq(0, 80, 5)) + 
  scale_y_continuous(limits = c(0, 4), breaks = seq(0, 4, 0.5)) +

  labs(x = "Age in years", y = "Average imprecision in target width")
```

>>>>

write.csv(mgp_triallevel, "../saves/lifespan_mgp_triallevel.csv", quote = F, row.names = F)
saveRDS(mgp_triallevel, "../saves/lifespan_mgp_triallevel.rds")
mgp_triallevel <- readRDS("../saves/lifespan_mgp_triallevel.rds")
```

@@

write.csv(mgp_aggregated, "../saves/lifespan_mgp_aggregated.csv", quote = F, row.names = F)
saveRDS(mgp_aggregated, "../saves/lifespan_mgp_aggregated.rds")
mgp_aggregated <- readRDS("../saves/lifespan_mgp_aggregated.rds")
```

@@

ggplot() +
  geom_jitter(
    data = mdata, 
    aes(x = age, y = imprecision), 
    col = "darkgrey", alpha = 0.1
  ) +

  geom_smooth(
    data = pred_combined, 
    aes(x = age, y = Estimate, ymin = Q2.5, ymax = Q97.5, group = model, col = model), 
    stat = "identity", size = 0.75, alpha = 0.15
  ) +

  scale_color_colorblind() + 
  scale_x_continuous(breaks = seq(0, 80, 5)) +
  coord_cartesian(ylim = c(0, 4)) +
  labs(x = "Age in years", y = "Imprecision in target width", col = "Model structure")

ggsave("../figures/supplements_lifespan_modelstructure.png", width = 7, height = 5, scale = 1, bg = "white")
ggsave("../figures/supplements_lifespan_modelstructure.pdf", width = 7, height = 5, scale = 1, bg = "white")
```

### 2023-10-04 supplements/supplements.Rmd
ready for first round of co-author feedback

  - name          : "Julia Prein"
    affiliation   : "1"
  - name          : "Manuel Bohn"
    affiliation   : "1,2"
  - name          : "Luke Maurits"
    affiliation   : "1"
  - name          : "Annika Werwach"
    affiliation   : "1"
  - name          : "Daniel B. M. Haun"
    affiliation   : "1"
affiliation:
  - id            : "1"
    institution   : "Department of Comparative Cultural Psychology, Max Planck Institute for Evolutionary Anthropology, Leipzig, Germany"
  - id            : "2"
    institution   : "Institute of Psychology, Leuphana University Lüneburg, Germany"  
floatsintext      : yes
bibliography      : ["../../references.bib"]
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no
documentclass     : "apa7"
classoption       : "man"
output            : papaja::apa6_word
header-includes:
  - \usepackage{setspace}
  - \captionsetup[figure]{font={stretch=1}}
editor_options: 
  markdown: 
    wrap: sentence
---

```{r setup, include = FALSE}
library("papaja")
library("tidyverse")
library("ggpubr")
library("kableExtra")
theme_set(theme_classic())
```

@@

# Components of gaze understanding
## Theory of mind battery

@@

## Model comparison

>>>>

  - name          : "Julia Prein"
    affiliation   : "1"
  - name          : "Luke Maurits"
    affiliation   : "1"
  - name          : "Annika Werwach"
    affiliation   : "1"
  - name          : "Daniel B. M. Haun"
    affiliation   : "1,*"
  - name          : "Manuel Bohn"
    affiliation   : "1,2,*"
affiliation:
  - id            : "1"
    institution   : "Department of Comparative Cultural Psychology, Max Planck Institute for Evolutionary Anthropology, Leipzig, Germany"
  - id            : "2"
    institution   : "Institute of Psychology, Leuphana University Lüneburg, Germany"
  - id            : "*"
    institution   : "shared last authorship"  
floatsintext      : yes
bibliography      : ["../../references.bib"]
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no
documentclass     : "apa7"
classoption       : "man"
output            : papaja::apa6_pdf
header-includes:
  - \usepackage{setspace}
  - \usepackage{float}
  - \captionsetup[figure]{font={stretch=1}}
editor_options: 
  markdown: 
    wrap: sentence
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.pos = 'H')
```

```{r setup, include = FALSE}
library("papaja")
library("tidyverse")
library("ggpubr")
library("kableExtra")
library("brms")
theme_set(theme_classic())
```

@@

# Study 1: Lifespan
## Participants
```{r lifespan_sample, include = T, out.width = "100%"}
lifespan_sample <- readRDS("../saves/lifespan_sample.rds")

lifespan_table <- lifespan_sample %>% 
  mutate(
    n = str_glue_data(., "{n_total} ({n_females} female)"), 
    `Age range` = str_glue_data(., "{age_min} - {age_max}"), 
  ) %>% 
  rename(
    `Age group` = age_group,
    `Age SD` = age_sd, 
    `Age mean` = age_mean
  ) %>% 
  select(`Age group`, n, `Age mean`, `Age range`, `Age SD`)

knitr::kable(lifespan_table, format = "latex", booktabs = TRUE) %>% 
  kable_styling(latex_options = c("scale_down", "HOLD_position"))
```

## Analysis
### Model structures
In the paper, we reported the following model structures: linear model: `mean_imprecision ~ age_centered`; quadratic model in `R: mean_imprecision ~ 1 + age_centered + I(age_centered^2)`; cubic model: `mean_imprecision ~ 1 + age_centered + I(age_centered^2) + I(age_centered^3);` Gaussian process model: `mean_imprecision ~ gp(age_centered, k=50, c=5/4, scale=TRUE)`. Note the additional parameters in the Gaussian process model. With the default settings, the underlying Gaussian process maths would get solved exactly. By providing the arguments `k` and `c`, we use an approximation process. The higher the value of k, the better the approximation: we have used `k=50` for faster processing speed and better diagnostics. `brms` suggests 5/4 as a value for `c`. Adding `scale=TRUE` is supposed to improve sampling speed and convergence.

Originally, we fitted the models on a trial-by-trial basis with the following structure in `R: performance ~ age + symmetricPosition + trialNr + (1 + symmetricPosition + trialNr | subjID)`. However, the Gaussian Process model was computationally heavy. Therefore, we simplified the model structure, aggregated data on a subject level, and included only age as an effect. We then visually compared the model predictions of the original and the simplified models with each other. As you can see in Figure \@ref(fig:fig1), results of the two models did not differ notably.

(ref:figlab1) **Comparison between models on a trial- or subject-level**. Grey dots show data of each trial. Solid lines show the mean predicted developmental trajectory for both models. Line color denotes model structure (yellow: trial-level; black: aggregated on subject-level). 
```{r fig1, include = T, out.width = "100%", fig.align = "center", fig.cap = "(ref:figlab1)"}
knitr::include_graphics("../figures/supplements_lifespan_modelstructure.pdf")
```

### Changepoint analysis
In our Bayesian changepoint analysis, we restricted the model to a constant mean (i.e., a flat line with zero degree polynomial) within each segment, and to have minimally 10 data points between two change points (i.e., corresponding to half of the data points we collected per adult decade) to avoid "overreactions" to individual outlying data points. In a supplementary analysis, we assessed how different parameter settings effected of our changepoint analysis. We changed the number of allowed change points, the minimum number of data points between change points, and the polynomial order. When the model had more explorative room, for example, by a greater number of change points, smaller minimum number of data points between change points, higher polynomial order, the model outputs showed more fine-grained change points. The exact location of the change points varied slightly. Overall, the interpretation stayed the same as the one we reported in the paper. While early childhood was characterized by much change, adults showed a relatively stable level of imprecision. There was a minor change in that elderly adults became slightly more imprecise again. If you are interested into the details, please have a look at the file `supplements_changepoint_parameters.html`, which you can find in the GitHub repository in the `stats` folder.

# Study 2: Computational cognitive model 
## Analysis
### Gaze model prediction
Our gaze model predicts that TANGO trials vary in their difficulty, resulting in a U-shaped pattern: Participants' imprecision should increase, the further out the target lands (towards the very left/right sides). Since the task is presented on a screen, there is a natural border towards one side. Imagine the target lands to the very right side. Participants' imprecision cannot click further right because the screen ends; all their uncertainty about the target location faces the inner, left-hand side now. Therefore, the predicted U-shaped pattern should decrease again towards the screen borders. For previous reliability analyses [@prein2023tango], we had increased the trial number for an adult sample (N = 70; each 30 trials). Interestingly, here we found the expected shape in the data: the U-pattern decreased again towards the screen ends (Figure \@ref(fig:fig2)).

(ref:figlab2) **Gaze funnel for adult sample with higher trial number**.
```{r fig2, include = T, out.width = "100%", fig.align = "center", fig.cap = "(ref:figlab2)"}
knitr::include_graphics("../figures/gazefunnel_adults.pdf")
```

## Inference parameter estimates per individual
As can be seen in Figure \@ref(fig:fig3) and Figure \@ref(fig:fig4), the gaze model estimated the inference parameter for each individual. Across individuals, the inference parameter varied in the estimated magnitude and level of uncertainty. In general, estimates for more precise individuals (i.e., smaller inference parameter value) showed decreased levels of uncertainty.

(ref:figlab3) **Gaze model estimates faceted by individual**.
```{r fig3, include = T, out.width = "100%", fig.align = "center", fig.cap = "(ref:figlab3)"}
knitr::include_graphics("../figures/supplements_gazemodel_facetedinference.pdf")
```

(ref:figlab4) **Gaze model estimates ordered by descending inference parameter value**.
```{r fig4, include = T, out.width = "100%", fig.align = "center", fig.cap = "(ref:figlab4)"}
knitr::include_graphics("../figures/supplements_gazemodel_indinference.pdf")
```

# Study 3: Components of gaze understanding
## Procedure Theory of mind battery

@@

## Analysis
### Animal vs. human faces

```{r animal_human}
mstimuli <- readRDS("../saves/supplements_magnet_mstimuli.rds")
```

In Study 1 and Study 2, we presented the TANGO [@prein2023tango] with animal characters. For Study 3, we exchanged the animals with human faces, modelled in appearance after the local population. We decided to do so in order to enhance the social context of this task and to make it more comparable to the Theory of Mind task battery (where there is live interaction with the experimenter). To ensure the change from animal to human faces did not notably change children's responses, we conducted an exploratory analysis. We conducted a GLMM analysis with the following model structure in `R: click ~ age_scaled + stimuli + symmetric_position + trial_nr + (1 + symmetric_position + trial_nr | subj_id)`; where `stimuli` denoted either human or animal faces. The estimate for the fixed effect of `stimuli` was small and the 95% CrI included zero: $\beta$ = `r fixef(mstimuli)["stimulihumans", "Estimate"]`; 95% CrI [`r fixef(mstimuli)["stimulihumans", "Q2.5"]`; `r fixef(mstimuli)["stimulihumans", "Q97.5"]`]). Therefore, we concluded that the animal vs. human version of the TANGO did not differ substantially.

### Model comparisons
To identify which (social-)cognitive components were needed to best explain the TANGO score, we compared GLMMs that predicted the mean imprecision in gaze understanding by age + the respective task score: imprecision in non-social vector estimation, the ToM aggregate score, and/ or the aggregate of the two perspective-taking tasks (subset of ToM battery). For example, the model notation in `R: tango_mean ~ age_centered + magnet_scaled + perspective_scaled`). The model including the non-social vector estimation task (magnet) and the two perspective-taking tasks won, as indicated by the model comparison results below.
